---
title: "Stat 451 - Homework 5"
author: 'Due: 2/10 Before Class'
# date: "2/2/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(R2jags)
```

### Introduction
BEFORE STARTING(!!!): REMEMBER THAT BAYESIANS USE THEIR PRIOR KNOWLEDGE TO INFLUENCE THEIR BELIEFS. Looking at the data to inform your priors is considered 'double-dipping' into the data. Before reading in the data, gather some intuition about the number of hot dogs that an AVERAGE person can eat in 1 minute (don't think too hard -- experiments welcome :) ) and the number of miles that an average student can run. These ideas will influence your priors for this homework.

### Data
In a wild turn of events, you are asked to provide some understanding about the number of hot dogs (per minute) that some amateur, hot-dog-eating BYU statistics students can scarf down. Data provided to you is the number of `hot dogs` (per minute) that these students can eat and the maximum number of `miles` that these students can run. Read in the data as:

```{r}
dat <- read.table("linregdata.dat", header=TRUE)
colnames(dat) <- c('contestant', 'miles_run', 'hotdogs_minute')
# dat
```


### Assignment
* 1. Write JAGS code to perform a Bayesian Linear Regression with `hot dogs per minute` as the response variable and `running miles` as the explanatory variable. 
  + 1.a Use your intuition gathered to set appropriate priors on the parameters $\beta_0$, $\beta_1$, and $\sigma^2$.
  + 1.b Assess convergence of parameters using trace plots and convergence diagnostics. Note in your answers what appropriate ranges are for the diagnostics and why you believe that your draws did/didn't converge.
```{r part1}
# linear model
mdl <- "
  model {

    for (i in 1:31) {
      y[i] ~ dnorm(mu[i], 1/vr)
      mu[i] <- b0 + b1*x[i]
    }
    b0 ~ dnorm(1, 10)
    b1 ~ dnorm(0.5, 1)
    vr ~ dgamma(1.1, 0.1)
  }
"


writeLines(mdl, 'linreg.txt')

x <- dat$miles_run
y <- dat$hotdogs_minute

data.jags <- c('x', 'y')
parms <- c('b0', 'b1', 'vr')

linreg.sim <- jags(data=data.jags, inits = NULL, parameters.to.save = parms,
                   model.file = 'linreg.txt', n.iter = 4000, n.burnin = 1000,
                   n.chains = 4, n.thin = 3)

linreg.sim
```

* 2. Plot a fitted values line (for reasonable values of miles run) on a scatterplot of your data.
  + 2.a Add confidence intervals around the line, denoting our uncertainty of $E(Y|X)$. Color and label the lines.
  + 2.b Add prediction intervals around the line, denoting our uncertainty of a future prediction. Color and label the lines.

* 3. Explain how Bayesian Linear Regression differs from Frequentist Liner Regressiona.

* 4. Joey Chestnut, the world record holder hot dog eater, can eat 7.6 hot dogs per minute. He can also run 7.25 miles. Does he fall outside the prediction intervals for a person who can run 7.25 miles? What is his error, $\epsilon_{Joey} = \hat{y}_{Joey} - y_{joey}$?





