---
title: "Homework 7"
author: 'Garrett Carr'
output: pdf_document
---

For this assignment, your task is simple: The researchers want to know which seed type, cultivar, and/or seed type by cultivar has the best yield. 

Use JAGS or Stan to perform this analysis. You will be graded on your ability to perform a complete analysis.

```{r setup, include=FALSE}
library(R2jags)
cult <- read.table("cult.dat")
colnames(cult) <- c('field', 'type', 'cultivar', 'yield')
head(cult)
cult$type <- as.factor(cult$type)
cult$cultivar <- as.factor(cult$cultivar)
```

This is a classic split plot, with a potential interaction affect. Since we are interested
 in whether seed type, cultivar, or a combination affects yield, I assumed that
 any affects introduced by differences between fields were not interesting.

```{r model1}
model1 <- "
model {
for (i in 1:24) {
  yield[i] ~ dnorm(mu[i], 1/s2error)
  mu[i] <- alpha0 + field[i] + seed[type[i]] + food[cultivar[i]] + interax[type[i], cultivar[i]]  
}
  alpha0 ~ dnorm(0, 0.000001)
  for (i in 1:4) {
    field[i] ~ dnorm(0, 0.01)
  }
  
  for (i in 1:2) {
    seed[i] ~ dnorm(30, 1/s2seed)
  }
  
  for (i in 1:3) {
    food[i] ~ dnorm(0, 1/s2cult)
  }
  
  for (i in 1:2) {
    for (j in 1:3) {
      interax[i,j] ~ dnorm(0, 0.001)
    }
  }
  s2error ~ dgamma(1.1, 0.05)
  s2cult ~ dgamma(1.1, 0.05)
  s2seed ~ dgamma(1.1, 0.05)
  
}
"

writeLines(model1, 'hw7model.txt')

field <- cult$field
type <- cult$type
cultivar <- cult$cultivar
yield <- cult$yield

data.jags <- c('field', 'type', 'cultivar', 'yield')
parms <- c('seed', 'food', 'interax', 's2error', 's2cult', 's2seed', 'alpha0')

cult.sim <- jags(model.file = 'hw7model.txt', data = data.jags, parameters.to.save = parms,
                 n.iter = 5000, n.burnin = 1000, n.chains = 4, n.thin = 1, inits = NULL)

cult.sim
```
It appears there is hardly a difference between seed one and two. 


```{r model2}
model2 <- "
model {
for (i in 1:24) {
  yield[i] ~ dnorm(mu[i], 1/s2error)
  mu[i] <- alpha0 + field[i] + seed[type[i]] + food[cultivar[i]] + interax[type[i], cultivar[i]]  
}
  alpha0 ~ dnorm(30, 0.0001)
  for (i in 1:4) {
    field[i] ~ dnorm(0, 0.01)
  }
  
  for (i in 1:2) {
    seed[i] ~ dnorm(0, 1/s2seed)
  }
  
  for (i in 1:3) {
    food[i] ~ dnorm(0, 1/s2cult)
  }
  
  for (i in 1:2) {
    for (j in 1:3) {
      interax[i,j] ~ dnorm(0, 0.01)
    }
  }
  s2error ~ dgamma(1.1, 0.05)
  s2cult ~ dgamma(1.1, 0.05)
  s2seed ~ dgamma(1.1, 0.05)
  
}
"

writeLines(model2, 'hw7model2.txt')

cult.sim <- jags(model.file = 'hw7model2.txt', data = data.jags, parameters.to.save = parms,
                 n.iter = 5000, n.burnin = 1000, n.chains = 4, n.thin = 1, inits = NULL)

cult.sim
```
Just based on what we are seeing so far, I would say that the best bet is to use seed type 2
with cultivation 3. So, type b, with liv as the food type.
