---
title: "HW4"
author: "Garrett Carr"
date: "2/1/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(R2jags)
library(tidyverse)
library(coda)
library(GGally)

oxidedat <- read.table('oxide.dat', header = TRUE)

oxidedat$type <- as_factor(oxidedat$type)
```

We read in the data and take a look:
```{r data}
head(oxidedat,14)
```
The factors being examined are where in the machine the wafers were located (`location`), and the type of wafer being used (`virgin` - a brand new wafer, `in-house`, a wafer refurbished by the company, and `ext` - a refurbished wafer purchased from an outside vendor). 

The variable being tested is the thickness of oxide ayer labeled `oxide`. The thicker the oxide layer the better.


Oxide layer appears to be normally distributed.

```{r EDA}
hist(oxidedat$oxide)

oxidedat %>% group_by(location) %>% summarise(.groups = 'keep', avg = mean(oxide), v = var(oxide))
oxidedat %>% group_by(type) %>% summarise(.groups = 'keep', avg = mean(oxide), v = var(oxide))
oxidedat %>% group_by(location, type) %>% summarise(.groups = 'keep', avg = mean(oxide), v = var(oxide))
```

It also appears there is probably some interaction based on the averages. 
It also appears that in general, there may be larger differences in variation between wafer types, compared to locations.

```{r model1}
mdl <- "
model {
  for (i in 1:27) {
    oxide[i] ~ dnorm(mu[location[i], type[i]], 1/vr[location[i],type[i]])
  }
  # Priors
  for (i in 1:3) {
    for (j in 1:3) {
      mu[i,j] ~ dnorm(90, 0.01)
      vr[i,j] ~ dgamma(2,1.65)
    }
  }
}
"

writeLines(mdl, 'oxide.txt')
oxide <- oxidedat$oxide
location <- oxidedat$location
type <- oxidedat$type
rep <- oxidedat$rep
data.jags <- c('oxide', 'location', 'type')
parms <- c('mu', 'vr')

oxide.sim <- jags(data = data.jags, inits = NULL, parameters.to.save = parms, 
                  model.file='oxide.txt', n.chains = 3, n.iter = 5000, 
                  n.burnin = 1000, n.thin = 4)

oxide.sim
```





```{r look}
sims <- as.mcmc(oxide.sim)
chains <- as.matrix(sims)
simsall <- as.mcmc(chains)
```

```{r convergence}
# plot(sims)

geweke.diag(sims)
geweke.diag(simsall)
```

It appears that vr[3,2] and vr[3,3] may have some issues, and may not have burned in enough.

I will increase my burn in.

```{r gelman}
gelman.diag(sims)
```
It appears that vr[3,3] seems to be off according to the gelman diagnostic. It probably didn't converge because I didn't do enough burn in.

```{r raferty}
raftery.diag(simsall)
```

According to raftery diagnostic, I need a higher sample size. 


Since I believe we don't have enough information to measure variance between all treatment types, I am only using the type to inform our model on variances, since those might seem to produce larger differences.

Here is the updated model:

```{r model2}
mdl <- "
model {
  for (i in 1:27) {
    oxide[i] ~ dnorm(mu[type[i], location[i]], 1/vr[type[i]])
  }
  # Priors
  for (i in 1:3) {
    for (j in 1:3) {
      mu[i,j] ~ dnorm(90, 0.01)
    }
    vr[i] ~ dgamma(6,0.5)
  }
}
"

writeLines(mdl, 'oxide.txt')
oxide <- oxidedat$oxide
location <- oxidedat$location
type <- oxidedat$type
rep <- oxidedat$rep
data.jags <- c('oxide', 'location', 'type')
parms <- c('mu', 'vr')

oxide.sim <- jags(data = data.jags, inits = NULL, parameters.to.save = parms, 
                  model.file='oxide.txt', n.chains = 3, n.iter = 11000, 
                  n.burnin = 2000, n.thin = 2)

oxide.sim
```
```{r look2}
sims <- as.mcmc(oxide.sim)
chains <- as.matrix(sims)
simsall <- as.mcmc(chains)

geweke.diag(sims)
geweke.diag(simsall)
```

This model appears to pass the geweke diagnostic. We check the gelman diagnostic:

```{r gelman2}
gelman.diag(sims)
```
Appears to pass.

```{r raftery2}
raftery.diag(sims)
raftery.diag(simsall)
```

raftery shows a pass, as each is below 5, mostly hovering around 2.

```{r plot2}

plot(chains[,2], type = 'l')
plot(chains[,8], type = 'l')
plot(chains[,9], type = 'l')
plot(chains[,10], type = 'l')
```

Traceplots check out. Lets look at the variances:

```{r var-traceplot}
plot(chains[,11], type = 'l')
plot(chains[,12], type = 'l')
plot(chains[,13], type = 'l')
```

```{r auto-corr}
autocorr.diag(simsall)
```
Autocorrelation appears to be good.
```{r eff-size}
effectiveSize(sims)
```
Everything is above 5000 effective sample sizes. I think this is sitting well, and we are ready for inference.

## Summary

```{r summary}
mu11 <- chains[,2]
mu12 <- chains[,3]
mu13 <- chains[,4]
mu21 <- chains[,5]
mu22 <- chains[,6]
mu23 <- chains[,7]
mu31 <- chains[,8]
mu32 <- chains[,9]
mu33 <- chains[,10]
vr1 <- chains[,11]
vr2 <- chains[,12]
vr3 <- chains[,13]
```

Looking for interactions:

```{r interactions}
int1 <- mu11-mu12-mu21+mu22
int2 <- mu12-mu13-mu22+mu23
int3 <- mu11-mu12-mu31+mu32
int4 <- mu12-mu13-mu32+mu33
int5 <- mu11-mu13-mu21+mu23
int6 <- mu21-mu22-mu31+mu32
int7 <- mu22-mu23-mu32+mu33
int8 <- mu21-mu23-mu31+mu33
int9 <- mu11-mu13-mu31+mu33

plot(density(int1))
plot(density(int2))
plot(density(int3))
plot(density(int4))
plot(density(int5))
plot(density(int6))
plot(density(int7))
plot(density(int8))
plot(density(int9))

mean(int1>0)
mean(int2>0)
mean(int3>0)
mean(int4>0)
mean(int5>0)
mean(int6>0)
mean(int7>0)
mean(int8>0)
mean(int9>0)
```

It appears there is an interaction for int6, and to a lesser extent, int4, 5, and 7.
Of these, int4 and 7 seem to be related to the strong interaction effect in mu32,
while 5 is probably a secondary effect.

Based on these, I would say that the strongest interaction effects are produced by location 2,
and wafer type 3 (ext).

```{r variance-analysis}
oxide.sim
```
Based on the above analysis, and the apparent negative interaction between the type 3
wafer and location 2, I would suggest that chip location 2 with in-house refurbished 
wafers will produce among the best results the most consistently. 

The analysis showed that type 2 wafers (in-house refurbished) are typically more 
consistent with oxide layer yields, with a lower overall standard deviation 
with a 90% probability interval between `r round(sqrt(quantile(chains[,12], c(0.1, 0.9))), 3)`.

