---
title: "Stat 251 Final Exam"
author: "FirstName LastName"
date: "12/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=4.5, fig.align='center')
```

## Problem [choose 1 or 2]

Chosen problem 1
a. 
```{r answera}
# Priors
lambda.p <- 0
tau2.p <- 1000
alpha.p <- 2
beta.p <- 20
```
We set our prior parameters.

    beta_0 ~ N(0, 1000)
    beta_1 ~ N(0, 1000)
    sigma^2 ~ Gamma(2, 20)
    
Since we know the full conditionals of $\beta_0$ and $\beta_1$, we use that and draw random samples from them.

```{r answe-a-cont}
x_i <- c(30.8, 26.4, 26.9, 27.9, 27.9, 26.8, 29, 29.8, 27.2, 29.7)
y_i <- c(0.516, 0.376, 0.37, 0.434, 0.427, 0.382, 0.458, 0.455, 0.386, 0.490)
n <- 10
```

We have to compute $\sigma^2$ first, since it's part of the full conditional of the betas.

```{r}

# Loglik function
loglik <- function(beta0, beta1, sigma2){
  sum(dnorm(y_i, beta0 + x_i*beta1, sqrt(sigma2), log=TRUE))
}

# Betas
log.beta0 <- function(sigma2val, beta1val) {
  dnorm((1000*sum(y_i - beta1val*x_i))/(1000*n + sigma2val), 
         sqrt(1000 * sigma2val/(1000*n + sigma2val)), 
         log=TRUE)
}

log.beta1 <- function(sigma2val, beta0val) {
  dnorm((1000*sum((y_i - beta0val)*x_i))/(1000*sum(x_i^2) + sigma2val), 
         sqrt(1000 * sigma2val/(1000*sum(x_i^2) + sigma2val)), 
         log=TRUE)
}

# Prior for sigma2

logprior.sigma2 <- function(sigma2val, alpha.p, beta.p) {
  dgamma(sigma2val, alpha.p, beta.p, log = TRUE)
}
```

```{r MCMC}
# MCMC setup!
iters <- 1000000
beta0.save <- rep(0, iters)
beta1.save <- rep(0, iters)
sigma2.save <- rep(0, iters)
beta0 <- -0.5
beta1 <- 0.03
sigma2 <- .0001
s.beta0 <- 0.005 # Step size
s.beta1 <- 0.0008 # Step size
s.sigma2 <- 0.0005
accept.beta0 <- 0
accept.beta1 <- 0
accept.sigma2 <- 0

# MCMC
for(t in 1:iters){
  # Sampling from sigma2 full conditional
  sigma2.dot <- rnorm(1, sigma2, s.sigma2) #proposed value
  if(sigma2.dot > 0){
    logr <- loglik(beta0, beta1, sigma2.dot) + logprior.sigma2(sigma2.dot, alpha.p, beta.p) -
      loglik(beta0, beta1, sigma2) - logprior.sigma2(sigma2, alpha.p, beta.p)
    logunif <- log(runif(1))
    if(logunif < logr){
      sigma2 <- sigma2.dot
      accept.sigma2 <- accept.sigma2 + 1
    }
  }
  sigma2.save[t] <- sigma2

  beta0.dot <- rnorm(1, beta0, s.beta0) # new proposed value for beta0
  logr <- loglik(beta0.dot, beta1, sigma2) + log.beta0(sigma2, beta1) - log.beta0(sigma2, beta1) - loglik(beta0, beta1, sigma2)
  logunif <- log(runif(1))
  if(logunif < logr){
    beta0 <- beta0.dot
    accept.beta0 <- accept.beta0 + 1
  }
  beta0.save[t] <- beta0

  beta1.dot <- rnorm(1, beta1, s.beta1) # new proposed value for beta1
  logr <- loglik(beta0, beta1.dot, sigma2) + log.beta1(sigma2, beta0) - log.beta1(sigma2, beta0) - loglik(beta0, beta1, sigma2)
  logunif <- log(runif(1))
  if(logunif < logr){
    beta1 <- beta1.dot
    accept.beta1 <- accept.beta1 + 1
  }
  beta1.save[t] <- beta1
}
```
Checking acceptance rates and trace plots.



b. 
```{r answerb}
accept.sigma2/iters
accept.beta0/iters
accept.beta1/iters

plot(sigma2.save, type = "l")
plot(beta0.save, type = "l")
plot(beta1.save, type = "l")
```
  

c. 
```{r answerc}
plot(density(beta1.save))
```


d. 
```{r answerd}
interval <- quantile(beta1.save, c(0.025, 0.975))
interval
```
We can say that for every 1 degree increase in temperature (F), we have an average increase in weight between `r round(interval[[1]], 3)` and `r round(interval[[2]], 3)` lbs per cubic foot, with 95% probability.

e. 
```{r answere}
# compute using drawn values and monte carlo approximation
# y_i = beta0.save + x_i*beta1.save + sqrt(sigma2.save) (error)

y <- rnorm(length(beta0.save), beta0.save + 25*beta1.save, sqrt(sigma2.save))

interval <- quantile(y, c(0.025, 0.975))
interval
```
On a randomly selected day with a temperature of 25 degrees F, based on our data and prior knowledge there is a 95% probability of the snow weight being between `r round(interval[[1]], 3)` and `r round(interval[[2]], 3)` lbs per cubic foot, with 95% probability.

