---
title: "Exam 2"
author: "Stat 451"
date: "3/10/2022"
output: pdf_document
---

This is a take home-exam. The exam is due at the beginning of class on Tuesday, March $15^{th}$. Please do not consult with any living person about this exam.

The first part of the exam uses the data file **mortality.dat**, and will be
worth 75% of the grade.  
 In the data set, the
first column is an identifier for the hospital. The second column is the number of operations
taking place in the neonatal unit over a one year period. The third column is the number of
operations that ended with the death of the patient.
The idea behind the study is to rank the performances of the hospitals involved. I 
assume you will do the problem using JAGS.  Please include relevant code with your answers.


```{r setup}
library(coda)
library(R2jags)
dat <- read.table('mortality.dat', header = TRUE)
dat
```


1. Presumably, an appropriate way to rank the hospitals would be to model the proportion
of operations that end in the death of the patient ($p_i$, where i indicates the $i^{th}$ hospital). The maximum likelihood estimates
for those proportions would simply be the number of deaths divided by the number of
operations. Compute the MLE for each hospital.

```{r part 1}
dat$p_i <- dat$deaths / dat$ops
dat
```


2. Now, perform a Bayesian analysis where the proportion of deaths is modeled as a fixed
effect.  That is, you should not use a hierarchical model.  Use a beta(1,1) prior for
the $p_i$'s.  The answer for this problem is just the standard JAGS output.

```{r part2}
model1 <- "
model{

# Liklihood
for (i in 1:11) {
  effect[i] ~ dbin(prop[i], ops[i])
}

# Priors
for (i in 1:11) {
  prop[i] ~ dbeta(1,1)
}
}
"
writeLines(model1, 'exam2mdl1.txt')

prop <- dat$p_i
ops <- dat$ops

data.jags <- c('prop', 'ops')
parms <- c('effect')

mort.sim <- jags(model.file = 'exam2mdl1.txt', data = data.jags, parameters.to.save = parms, n.iter = 10000, n.burnin = 1000, n.chains = 4, n.thin = 1, inits = NULL)

mort.sim
```



3. Demonstrate that the parameters you are estimating in this model have converged.  Show
effective sample sizes, autocorrelation, and the Raftery-Lewis diagnostic.

```{r diagnostic3}
sims <- as.mcmc(mort.sim)
chains <- as.matrix(sims)
sims <- as.mcmc(chains)
effectiveSize(sims)
autocorr.diag(sims)
raftery.diag(sims)
```
Chains show convergence, as effective size is greater than 5000, autocorrelation is low,
and raftery lewis diagnostic shows there were enough samples.


4.  Now perform a Bayesian analysis where the proportion of deaths in modeled as a
random effect.  That is, now use a hierarchical model for the $p_i$. Use a beta($\alpha$,$\beta$) prior for the $p_i$'s and gamma(1.1,1.1) for $\alpha$ and $\beta$.  The answer for this problem is
just the standard JAGS output.

```{r part4}
model2 <- "
model{

# Liklihood
for (i in 2:11) {
  effect[i] ~ dbin(prop[i], ops[i])
}

# Priors
for (i in 2:11) {
  prop[i] ~ dbeta(alpha, beta) 
}
alpha ~ dgamma(1.1, 1.1)
beta ~ dgamma(1.1, 1.1)
}
"
writeLines(model2, 'exam2mdl2.txt')

deaths <- dat$deaths


data.jags <- c('prop', 'ops')
parms <- c('effect')

mort.sim2 <- jags(model.file = 'exam2mdl2.txt', data = data.jags, parameters.to.save = parms, n.iter = 5000, n.burnin = 1000, n.chains = 4, n.thin = 1)

mort.sim2
```
There was a problem, because the proportion provided for hospital 1 is 0, since they have so few
ops. The sampler doesn't like that, because there isn't any information on the likelihood for hospital 1,
as there is no demonstrated track record, and the priors are uninformative. The estimate would be 0.

Therefore, I only produced posteriors for hospitals 2-11, in order to follow the directions.

5.  Demonstrate that the parameters you are estimating in this model have converged.  Show
effective sample sizes, autocorrelation, and the Raftery-Lewis diagnostic.

```{r diag5}
sims2 <- as.mcmc(mort.sim2)
chains2 <- as.matrix(sims2)
sims2 <- as.mcmc(chains2)
effectiveSize(sims2)
autocorr.diag(sims2)
raftery.diag(sims2)
```

Diagnostics looks good.


6.  Produce a table with the MLE’s, the point estimates from the fixed effects model, and
the point estimates from the random effects model side-by-side.

```{r point-estimates}
#         MLE       , model1,           model2
hos1 <- c(dat$p_i[1], mean(chains[,2]), 0)
hos2 <- c(dat$p_i[2], mean(chains[,5]), mean(chains2[,4]))
hos3 <- c(dat$p_i[3], mean(chains[,6]), mean(chains2[,5]))
hos4 <- c(dat$p_i[4], mean(chains[,7]), mean(chains2[,6]))
hos5 <- c(dat$p_i[5], mean(chains[,8]), mean(chains2[,7]))
hos6 <- c(dat$p_i[6], mean(chains[,9]), mean(chains2[,8]))
hos7 <- c(dat$p_i[7], mean(chains[,10]), mean(chains2[,9]))
hos8 <- c(dat$p_i[8], mean(chains[,11]), mean(chains2[,10]))
hos9 <- c(dat$p_i[9], mean(chains[,12]), mean(chains2[,11]))
hos10 <- c(dat$p_i[10], mean(chains[,3]), mean(chains2[,2]))
hos11 <- c(dat$p_i[11], mean(chains[,4]), mean(chains2[,3]))

df <- as.data.frame(rbind(hos1, hos2, hos3, hos4, hos5, hos6, hos7, hos8, hos9, hos10, hos11))
colnames(df) <- c('MLE', 'fixed', 'random')

df$fixed <- df$fixed/dat$ops
df$random <- df$random/dat$ops
df
```


7. Using the point estimates, which hospital seems to be the worst?

It looks like hospital 8 is the worst.

8. Using the point estimates, which hospital seems to be the second worst?

It looks like hospital 11 is the second worst.

9. Using the random effects model, would you say those two hospitals have significantly different proportions or neonatal mortalities at the 0.05 level?

```{r}
# Hospital 8
HPDinterval(sims2[,10], 0.95)

# Hospital 11
HPDinterval(sims2[,3], 0.95)
```
It appears the only difference is one more death, so I would say the proportions are not significantly different. Hospital 11 just has more operations.


10. Using the random effects model, which of these two hospitals would you prefer for your own family for neonatal
surgery?

I would probably pick Hospital 11 out of the two, because they have more experience doing the operations.


11. (Bonus)  Using the random effects model, rank the hospitals by proportion of deaths (lowest proportion has rank 1) at each draw. Then produce a mean rank and standard deviation of that rank for each hospital.

For the last set of problems, you have been brought data by an anthropologist.  She has found five adult skeletons of ancient 
humanoids.  She has been studying the ratio of the length of humerus (upper arm bone) to the length of the 
femur (thigh bone) in primates.  For monkeys she knows that ratio is about 0.95.  For modern man, the ratio is about 0.72.
The ratios for the five adult skeletons she has found are as follows: 0.857, 0.824, 0.820, 0.875, 0.844.  She 
is interested in the probability that the population mean ratio for the group of people whose skeletons she has 
found is between 0.80 and 0.90.  For your likelihood, you should be aware that the appropriate support for these
data is between 0 and 1. So we will use a beta likelihood.  You should also know that these ratios range from about 0.70 to 0.98.  You can use that information to inform your prior.
The parameters of a beta must be positive, so we will use gamma priors for the parameters.  If the beta is
parameterized with ($\alpha$,$\beta$), then use a gamma(shape=1,rate=.2) for $\alpha$, and a gamma(shape=1.5,rate=1) for $\beta$.

12. Since you know the likelihood and the prior distributions for the parameters, you can draw values from the
prior predictive.  Draw 1000 values from the prior predictive and plot a histogram of the values
drawn from the prior predictive.  Remember that the values
must be between 0 and 1.

```{r part12, eval=FALSE}

# Liklihood
p[i] ~ dbeta(alpha, beta)

# Priors
alpha ~ dgamma(1, 0.2)
beta ~ dgamma(1.5, 1)
}
```

```{r part12b}
prior <- rbeta(1000, rgamma(1000, 1, 0.2), rgamma(1000, 1.5, 1))
hist(prior)
```


13. Write code in Stan to address the problem.  Use 20000 burnin and 100000 iterations, thin by 10 and produce 5 chains.  This will
give you 40000 MCMC draws of the posterior.  Print the summary of the simulation. Also include your Stan code in the R script as comments. This will make it easier for the grader to identify any mistakes and give feedback. 


```{r part13}
library(cmdstanr)

data <- c(0.857, 0.824, 0.820, 0.875, 0.844)

"data {
	int N;
	int ratio[N];
     }
parameters {

           }
model {

      }
generated quantities {
  vector[N] log_lik;
}"
```

14.  Verify that the chains you have produced have converged appropriately and have enough information to use to make inference by examining (and reporting) the effective sample size.  Effective sample sizes should exceed 5000.

15. Verify that the chains you have produced have converged appropriately and have enough information to use to make inference by examining (and reporting) the Raftery-Louis diagnostic.  Raftery-Louis diagnostics should be less than 3.

16.  Plot the posterior density of the mean ratio which would be computed as $\alpha$/($\alpha$+$\beta$).  

17. What is the 
probability the ratio for the population mean given the data from these five skeletons is between 0.80 and 0.90?

